Microplatforms: Product platforms as code

### 2018-06-20 paper.tex
add ref architecture

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[reprint,amsmath,amssymb,aps]{revtex4-1}

@@

Consider the diagram from figure \ref{fig:automationn} again, this time with infrastructure approaches and technologies superimposed on the spectrum (figure \ref{fig:automation-techniques}). With the currently available level of automation, it is now entirely possible for a cross-functional team of ten to operate a fully-featured product platform for their application. This level of vertical integration has numerous benefits.

@@

Targeting a cluster orchestration technology, applications become fully portable between any infrastructure supporting the technology from a developer’s laptop through any public cloud provider to on-premise solutions. In the past year, Kubernetes [] has rapidly become a de-facto standard abstraction in cluster orchestration: it is possible to run a local cluster using a tool like Minikube [], clusters can be provisioned using simple compute resources in different clouds with Kops [], all three major cloud providers are offering hosted services (GKE [], EKS [] and AKS []) and managed on-premise solutions also exist, e.g. OpenShift by RedHat []or Pivotal Kubernetes Service [].

@@

Immutability is typically quite easily attained by content addressing - referring to any data by a cryptographic hash of the content itself. This concept appears a few times in our approach: code stored in git is referred to by the sha1 hash of the contents of a given revision [], Docker images and layers can be immutably referred to by a \texttt{sha256} hash of their content [].

@@

\newpage
\clearpage
\appendix
\section{\label{sec:architecture}Microplatform Reference Architecture using Kubernetes}

A microplatform reference architecture should describe everything that is required to be able to create and configure identical microplatforms, on any provider (including laptops, on-prem and cloud) in minimal time and with minimal effort (i.e. as fully automated as is possible). This includes sample code that actually works and all the associated documentation needed to quickly grok the system, so that an individual, who may be new to both the concepts and the implementation, can get immediate value.

A reference architecture does not represent a fully evolved or production-ready state, but may well be used to model an implementation that does.

@@

\end{document}
%
% ****** End of file apssamp.tex ******

>>>>

% ****** Start of file paper.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex paper.tex
%  2)  bibtex paper
%  3)  latex paper.tex
%  4)  latex paper.tex
%
\documentclass[reprint,amsmath,amssymb,aps]{revtex4-1}

@@

Consider the diagram from figure~\ref{fig:automationn} again, this time with infrastructure approaches and technologies superimposed on the spectrum (figure \ref{fig:automation-techniques}). With the currently available level of automation, it is now entirely possible for a cross-functional team of ten to operate a fully-featured product platform for their application. This level of vertical integration has numerous benefits.

@@

Targeting a cluster orchestration technology, applications become fully portable between any infrastructure supporting the technology from a developer’s laptop through any public cloud provider to on-premise solutions. In the past year, Kubernetes [] has rapidly become a de-facto standard abstraction in cluster orchestration: it is possible to run a local cluster using a tool like Minikube~\cite{RunningK70:online}, clusters can be provisioned using simple compute resources in different clouds with Kops [], all three major cloud providers are offering hosted services (GKE [], EKS [] and AKS []) and managed on-premise solutions also exist, e.g. OpenShift by RedHat []or Pivotal Kubernetes Service [].

@@

Immutability is typically quite easily attained by content addressing~- referring to any data by a cryptographic hash of the content itself. This concept appears a few times in our approach: code stored in git is referred to by the \texttt{sha1} hash of the contents of a given revision [], Docker images and layers can be immutably referred to by a \texttt{sha256} hash of their content [].

@@

\newpage
\clearpage

\appendix
\section{\label{sec:architecture}Microplatform Reference Architecture using Kubernetes}

\begin{figure*}[t!]
  \includegraphics[width=0.9\textwidth]{figs/workflow}
	\caption{Build, test, and deployment workflow}
	\label{fig:workflow}
\end{figure*}

A microplatform reference architecture should uphold all of the principles discussed above: full automation, immutability, idempotency, autonomy, and most of all, simplicity. It should use techniques such as Everything-as-Code, with declarative (rather than imperative) code where possible, which describes the desired state and is stored in a version controlled repository for audit and, most importantly, the ability to evolve. This results in Agile applications, built using an Agile architecture, running on Agile infrastructure.

This section describes a reference architecture that attempts to meet these criteria, but does not represent a fully evolved or production-ready state, although it may well be used to model an implementation that does.

We use a “forward-only” paradigm throughout. This means that every change is first written as code, committed to a git repository, a pull request is raised and peer reviewed, ultimately triggering a re-creation of the relevant resources. This applies equally to the infrastructure, architecture and the applications themselves.

We’ll first look at the flow of application build, test, and deploy, from an engineer’s laptop through to production. Then we’ll look at the cluster configuration and cloud architecture.

\subsection{\label{sec:buildflow}Build, test and deployment flow}

Figure~\ref{fig:workflow} shows the forward-only flow of changes, all the way to production. This shows how Continuous Integration (CI) to a monorepo triggers Continuous Deployment to Production (CDP), or, if there are human gates, Continuous Delivery (CD) into each environment.

\subsection{\label{sec:initialchange}It starts with a change to a component}

A developer makes a change to one or more of the application’s microservices (or its architecture, or infrastructure), builds and tests the change locally (not shown in the figure, but using an identical cluster configured in Minikube~\cite{RunningK70:online}) and then pushes the change to a monorepo. It forms part of a Pull Request, which is tested and peer reviewed before merging to master. While it is still on a branch, though, the push is communicated to a bot in the tooling cluster (CI) via a webhook. The bot triggers a build and test workflow, which in this diagram is shown as being implemented in Argo~\cite{argoproj96:online} because, in our opinion, this is a well implemented workflow engine that is well suited for running CI Docker builds inside containers running in a Kubernetes cluster. Because it is Kubernetes-native, Argo gives us the ability to leverage the declarative nature of Kubernetes resources, including Custom Resource Definitions (CRDs). This in turn allows us to store the workflow configuration, and the full CI/CD pipeline, as code in the monorepo alongside the application microservices. (Argo is much lighter than Jenkins~\cite{Jenkins9:online} for example and gives us full control of every component’s build and test - each step in the workflow is entirely declared as code in the monorepo~\cite{Monorepo54:online} and is run in its own Kubernetes pod).

This is Pipeline-as-Code and keeps the build and deploy in step with the application, its architecture and infrastructure.

\subsection{\label{sec:monorepo}A note about the monorepo}

\begin{figure*}[t!]
	\includegraphics[width=0.9\textwidth]{figs/monorepo}
	\caption{Atomic commits across a monorepo}
	\label{fig:monorepo}
\end{figure*}

It’s important to use a monorepo so that we can do atomic commits across the whole application (including the pipeline). Figure~\ref{fig:monorepo} shows how we can use the git commit SHA (e.g. db77f39) as a version number. This version represents a single handle on our application which is consistent across all microservices and associated artefacts. It’s a pointer to a configuration where every component is designed, built and tested to work together with every other component. It eliminates the need for SemVer~\cite{Semantic13:online} and massively reduces the complexity of dependency management within large microservice orchestrations, ultimately reducing risk and improving reliability of our application.

\subsection{\label{sec:dynamicworkflow}Dynamically configured workflow}

The triggered build and test workflow, first performs analysis of the push to the monorepo in order to identify which components have changed. A useful open source tool for this (written by the authors) is Monobuild~\cite{charypar24:online} which can produce a build schedule based on the hard (and soft) dependencies of changed components. This minimises the amount of work that is even attempted by the CI workflow and can be used to easily configure a dynamic Argo workflow template. However, because each component build is idempotent, building an unchanged component will result in the exact same artefacts, so that subsequent idempotent steps will also not result in anything being updated (or deployed).

\subsection{\label{sec:cba}Content based addressing}

Let’s pause for a moment to understand how content based addressing is important here. Git uses a Merkle tree~\cite{Merkletr83:online} where every commit is identified by a SHA~\cite{SecureHa13:online} (or hash) of its changes and those of every commit before it. We referred to using this as the version number in the section, above, on the monorepo. This SHA is a content-based address derived from the source code itself, where the same source code, with the same history, will produce the same SHA. This uniquely identifies every aspect of the source code itself and allows us to introduce idempotency into the build, where building the same commit will produce the same artefacts, which, importantly, are also identified using content based addressing.

If the artefact is a docker image, for example, then each layer is identified by a SHA of its bytes, and an image can be identified by a hash of all the layers’ hashes. This is used by the registry, to store the layers, and can be used by our pipeline during the deployment phases to ensure idempotency (more on this below).

In summary, content based addressing gives us idempotency. Builds are idempotent because of the source code hash, and registry pushes and service deployments are idempotent because of the hashes of the artefact binaries. All this means that we can actually submit a description of the entire service orchestration to the CI workflow and only the changed components would be rebuilt (only attempted if the source hash changed, and even then only the top layers of the container are likely to be rebuilt due to docker layer caching), pushed to the registry (also due to layer hashes), and deployed to the environment (due to the image hash).

\subsection{\label{sec:branchbuild}Branch build and test}

So the workflow template is populated for the changed components, and builds in series and in parallel based on the build schedule from monobuild. Unit tests are run as part of the Docker build, so an image is not created if any of the unit tests fail. Code quality analysis may also happen during the build. Docker multi-stage builds are used to ensure that the final image does not contain any of the build/test dependencies (i.e. the release artefacts are copied into a vanilla, lightweight container image that is usually based on alpine linux~\cite{indexAlp12:online}). This allows our release images to be much smaller, facilitating easier deployment and start up time, whilst also reducing our attack surface, as only software that is required at run time is included.

Each component is built and tested in its own Kubernetes pod, using a Docker-in-Docker~\cite{libraryd61:online} sidecar container running a Docker server. This means that layer caches from previous builds are not available, potentially meaning that work has to be re-done in order to successfully build a full image. To remediate this, the authors have written an open source drop-in replacement for “docker build”, called build-with-cache~\cite{redbadge83:online}. This small utility analyses the multi-stage Dockerfile to pull (and push back) images from previous builds of each target, allowing the current build to re-use unchanged layers. We have found that this can save a lot of unnecessary rework and reduce the build time considerably.

Successfully built images are pushed to a local registry, tagged with the commit SHA, ready for deployment to the dev cluster. Because the tag is the hash of the whole monorepo, an unchanged component will receive a new tag. This is not a problem, as we’ll see in the next section, because we resolve this to a hash of the image itself before deployment.

\subsection{\label{sec:branchdeploy}Branch deploy}

Outside in testing (such as BDD style functional tests and integration tests) need a running instance to test against. So, for every branch with an open pull request (PR), an instance of the whole application is deployed to the dev environment.

In order to ensure idempotency during deployment, the Kubernetes manifests (yaml files) that reference Docker images (e.g. deployment manifests) have their image names resolved before deployment. This is done by making a HEAD request to the registry and getting the image content digest from the response headers. This is important, because of the monorepo. Remember that there is only one commit SHA across all the components, so even if a component has not been changed, its SHA (and hence tag in the registry) will be updated. The resolution from commit SHA to content digest SHA brings this back in line, meaning that only manifests of changed components will end up being different from those of a currently deployed instance.

A URL (that includes the branch name) is posted back to the PR, so that developers, testers and stakeholders have easy access to review the new feature. The pipeline can now execute these tests against the instance, and report a successful (or failed) build back to the PR comments.

\subsection{\label{sec:gitops}GitOps}

In the reference architecture, deployment to each environment, is done through an associated git repository. This is called GitOps~\cite{GitOpsPa62:online}. We have written an open source tool (simply called deploy~\cite{redbadge45:online}) that can both raise pull requests against a cluster specific repo, and act as an agent (or bot) inside the associated target cluster that listens on a webhook and validates and deploys the updated applications to the cluster. It runs under a specified RBAC service account within the cluster, meaning that the CI workflow does not need elevated permissions in order to deploy to higher environments. GitOps allows for your configuration to be pulled by the cluster, rather than pushed by the CI system, and provides a point for controls to be implemented. The master branch of the cluster repo provides an non repudiable history of the cluster’s configuration, which is great for audit. Furthermore, historical configurations can be deployed to another identical cluster for fault finding. GitOps fits nicely into the DevSecOps~\cite{devsecop73:online} philosophy.

\begin{figure*}[t!]
	\includegraphics[width=0.9\textwidth]{figs/infrastructure}
	\caption{Cloud architecture with connection to on-prem backends}
	\label{fig:infrastructure}
\end{figure*}

\subsection{\label{sec:masterbuild}Master build, test and deploy}

When a pull request is merged to master, the exact same steps as described above are performed. The only difference is that these builds have the potential to progress further, through the environments, into production. Therefore, on successful completion of the build and test phases, the resulting images are pushed into a release candidate registry.

There may be human, or other control gates, in this part of the pipeline, but what’s important is that it is still 100% automated, and each environment deployment is identical to the one before it (with the exception that the images come from a release candidate registry). These deployments are also performed according to the GitOps philosophy, through the git repo that is associated with that environment’s cluster.

As the application is deployed to each higher environment, confidence in the correctness of the software is increased. Smoke tests in each environment are triggered by the workflow to ensure that the application has bedded into the environment correctly (e.g. it has discovered and connected to the right backends).

If CDP (rather than CD) is being practised, this final part of the workflow ensures that every commit to master will end up in production, if it passes all its tests and controls and has successfully been through all the lower environments, it will get to production. New features should be behind feature flags to allow separation of deployment from release.

\subsection{\label{sec:infrastructure}Infrastructure}

Infrastructure and cluster provisioning are fully automated, using tools such as Kops~\cite{kubernet77:online} and Terraform~\cite{Terrafor77:online}. Figure~\ref{fig:infrastructure}, shows cloud architecture where connection to on-prem backends is required.

First let’s look at the significant features of this architecture.

The Kubernetes clusters are provisioned in non-routable VPCs. There may be several of these VPCs, each with one or more clusters (with each cluster in its own private subnet). Each non-routable VPC has its own private address range, and these ranges can each be the same (e.g. 10.0.0.0/16), which means they don’t have to be assigned/managed and it gets us closer to our goal of having every environment be identical to every other. The fact that it is non-routable means that the private links (layer 4 routers with associated L4 load balancers) are the only way into and out of the VPC. These well-defined ingress and egress points provide control and mean that even if the VPC is compromised, the attacker would only be able to route traffic to known locations.

For high availability, masters, nodes, and proxies are provisioned across multiple availability zones. So, for instance, there would be 3 inner and 3 outer proxies, and at least 3 masters and and least 3 cluster nodes, spread across 3 availability zones.

Internet traffic comes in through an edge Web Application Firewall and Content Delivery Network, an API Gateway and a L7 load balancer so there are well defined opportunities to regulate incoming HTTP traffic.

All outbound traffic from the cluster must exit through a private link to an “inner” L7 forward proxy (e.g. squid), which routes Internet traffic through another private link to an “outer” L7 forward proxy in another non-routable (sidecar) VPC, and on-prem traffic through firewalled direct connect to internal destinations.

Only non-routable VPCs have Internet gateways attached, providing ingress-only to the cluster VPCs and egress-only from the sidecar VPC.

Ideally humans should not require access to any machines in this infrastructure because the infrastructure is immutable (can only be created and destroyed, not modified), and so it should be possible to deny access for ssh throughout (not only is port 22 not configured in any security groups, but the virtual machines do not even need to run an ssh daemon). If there is a problem somewhere, it is picked up by extensive monitoring and logging, the problem is fixed in the source code and the infrastructure reprovisioned.

@@

\end{document}
%
% ****** End of file paper.tex ******
